PANCHANGE_analysis

### 2020-08-20 PANCHANGE_analysis.Rmd
pcl summary

@@

```{r gather to long format Disorder}
dat.long.psych <- dat %>%
  gather(
    key = "Disorder_raw",
    value = "Diagnosis",
    all_of(disorder_columns)) %>%
  select(
    ID,
    Sample,
    Gender,
    Ethnicity,
    all_of(diff_score_cols),
    Disorder_raw,
    Diagnosis
  ) %>%
  mutate(Disorder =
           recode_factor(
             Disorder_raw,
             anxiety_disorders_numeric = "Anxiety disorders",
             depressive_disorders_numeric = "Depressive disorders",
             eating_disorders_numeric = "Eating disorders",
             obsessive_compulsive_disorders_numeric = "OCDs",
             psychotic_disorders_numeric = "Psychotic disorders",
             mhd.mania_hypomania_bipolar_or_manicdepression_numeric = "Bipolar disorders",
             mhd.posttraumatic_stress_disorder_ptsd_numeric = "PTSD",
             mhd.autism_aspergers_or_autistic_spectrum_disorder_numeric = "Autism",
             mhd.attention_deficit_hyperactivity_disorder_numeric = "ADHD",
             mhd.personality_disorder_numeric = "Personality disorders",
             control_numeric = "Controls"
           ))

@@

```{r Disorder by Gender}
Disorder_by_Gender <- dat.long.psych %>%
  filter(Diagnosis == 1) %>%
  count(Disorder, Gender) %>%
  mutate(Prop = round(n/sum(n), 2))

Disorder_by_Gender
```

```{r Disorder by Gender}
Disorder_by_Gender <- dat.long.psych %>%
  group_by(Disorder) %>%
  filter(Diagnosis == 1) %>%
  count(Gender) %>%
  mutate(Prop = round(n/sum(n), 2))

Disorder_by_Gender
```

```{r Disorder by Gender plot}
Disorder_by_Gender %>%
  ggplot(mapping = aes(
    x = Disorder,
    y = n)
    ) +
  geom_bar(
    aes(fill = Gender),
#    position = "dodge",
    stat="identity",
    binwidth = 1,
    position = position_dodge2(preserve = "single")
  ) +
  labs(
    x = "Disorder",
    y = "Number of participants",
    fill = "Gender") +
  theme_personal +
  scale_fill_manual(
    values = COPINGpalette4,
#    labels = c("Prepandemic", "Retrospective")
    ) +
  coord_flip()
```

```{r Disorder by Sample}
Disorder_by_Sample <- dat.long.psych %>%
  filter(Diagnosis == 1) %>%
  group_by(Disorder, Sample) %>%
  count(Disorder)
```

```{r Disorder by Sample plot}
Disorder_by_Sample %>%
  ggplot(mapping = aes(
    x = Disorder,
    y = n)
    ) +
  geom_bar(
    aes(fill = Sample),
#    position = "dodge",
    stat="identity",
    binwidth = 1,
    position = position_dodge2(preserve = "single")
  ) +
  labs(
    x = "Disorder",
    y = "Number of participants",
    fill = "Sample") +
  theme_personal +
  scale_fill_manual(
    values = COPINGpalette4,
#    labels = c("Prepandemic", "Retrospective")
    ) +
  coord_flip()
```

Ethnicity should be recoded
European
Mixed or multiple ethnic origins
African or African British
Asian or Asian British
Arab
```{r Disorder by Ethnicity}
Disorder_by_Ethnicity <- dat.long.psych %>%
  filter(Diagnosis == 1) %>%
  count(Disorder, Ethnicity) %>%
  mutate(Prop = round(n/sum(n), 2))

Disorder_by_Ethnicity
```

```{r Disorder by Ethnicity plot}
Disorder_by_Ethnicity %>%
  ggplot(mapping = aes(
    x = Disorder,
    y = n)
  ) +
  geom_bar(
    aes(fill = Ethnicity),
    #    position = "dodge",
    stat="identity",
    binwidth = 1,
    position = position_dodge2(preserve = "single")
  ) +
  labs(
    x = "Disorder",
    y = "Number of participants",
    fill = "Ethnicity") +
  theme_personal +
  scale_fill_manual(
    values = COPINGpalette6,
    #    labels = c("Prepandemic", "Retrospective")
  ) +
  coord_flip()
```

@@

```{r}
gad.diff_score_prepan_base_plot <- dat %>%
  ggplot(mapping = aes(y = gad.diff_score_prepan_base)) +
  geom_bar() +
  theme_minimal()

@@

```{r fig.height=5, fig.width=5}
gad.diff_score_prepan_base_plot /
  gad.diff_score_prepan_retro_plot /
  gad.diff_score_prepan_base_plot +
  plot_layout(guides = "collect", heights = c(1,1,1)) +
  plot_annotation(tag_levels = "A")
```

@@

```{r}
dat.long %>%
  filter(Diagnosis == 1) %>%
  group_by(Disorder) %>%
  summarise(
    gad_base_retro = mean(gad.diff_score_base_retro, na.rm = T),
    gad_prepan_retro = mean(gad.diff_score_prepan_retro, na.rm = T),
    gad_prepan_base = mean(gad.diff_score_prepan_base, na.rm = T),
    phq_base_retro = mean(phq.diff_score_base_retro, na.rm = T),
    phq_prepan_retro = mean(phq.diff_score_prepan_retro, na.rm = T),
    phq_prepan_base = mean(phq.diff_score_prepan_base, na.rm = T),
    ocir_base_retro = mean(ocir.diff_score_base_retro, na.rm = T),
    ocir_prepan_retro = mean(ocir.diff_score_prepan_retro, na.rm = T),
    ocir_prepan_base = mean(ocir.diff_score_prepan_base, na.rm = T)
    )
```

@@

Model 0: Disorders
```{r}
model0 <- lm(
  formula = gad.diff_score_base_retro ~ Disorder,
  data = dat.long)

@@

```{r}
model1 <- lm(
  formula = gad.diff_score_base_retro ~ time_diff_sign_up_coping,
  data = dat)

@@

Model 1: Model 0 + Sex + Age + Ethnicity + Time registration and baseline
```{r}
model1 <- lm(
  formula = gad.diff_score_base_retro ~ Gender + Age + Ethnicity + time_diff_sign_up_coping + Sample,
  data = dat)

Re-calculate change
```{r}

Re-rank the change again
```{r}

PCL:  
Calculate the total GLAD, COPING, RAMP baseline 
```{r}

>>>>

# Psychiatric disorder hierarchy

control_numeric == 1 ~ 0,
depressive_disorders_numeric == 1 ~ 1,
anxiety_disorders_numeric == 1 ~ 2,
mhd.posttraumatic_stress_disorder_ptsd_numeric == 1 ~ 3,
eating_disorders_numeric == 1 ~ 4,
obsessive_compulsive_disorders_numeric == 1 ~ 5,
mhd.attention_deficit_hyperactivity_disorder_numeric == 1 ~ 6,
mhd.personality_disorder_numeric == 1 ~ 7,
mhd.autism_aspergers_or_autistic_spectrum_disorder_numeric == 1 ~ 8,
mhd.mania_hypomania_bipolar_or_manicdepression_numeric == 1 ~ 9,
psychotic_disorders_numeric == 1 ~ 10

psychotic_disorders_numeric == 1 ~ 10
mhd.mania_hypomania_bipolar_or_manicdepression_numeric == 1 ~ 9,
mhd.autism_aspergers_or_autistic_spectrum_disorder_numeric == 1 ~ 8,
mhd.personality_disorder_numeric == 1 ~ 7,
mhd.attention_deficit_hyperactivity_disorder_numeric == 1 ~ 6,
obsessive_compulsive_disorders_numeric == 1 ~ 5,
eating_disorders_numeric == 1 ~ 4,
mhd.posttraumatic_stress_disorder_ptsd_numeric == 1 ~ 3,
anxiety_disorders_numeric == 1 ~ 2,
depressive_disorders_numeric == 1 ~ 1,
control_numeric == 1 ~ 0,

```{r Psychiatric disorders hierarchy}
dat <- dat %>%
  mutate(
    Disorder_numeric =
      case_when(
        psychotic_disorders_numeric == 1 ~ 11,
        mhd.mania_hypomania_bipolar_or_manicdepression_numeric == 1 ~ 10,
        mhd.autism_aspergers_or_autistic_spectrum_disorder_numeric == 1 ~ 9,
        mhd.personality_disorder_numeric == 1 ~ 8,
        mhd.attention_deficit_hyperactivity_disorder_numeric == 1 ~ 7,
        obsessive_compulsive_disorders_numeric == 1 ~ 6,
        eating_disorders_numeric == 1 ~ 5,
        mhd.posttraumatic_stress_disorder_ptsd_numeric == 1 ~ 4,
        (anxiety_disorders_numeric == 1 &
        depressive_disorders_numeric == 1) ~ 3,
        anxiety_disorders_numeric == 1 ~ 2,
        depressive_disorders_numeric == 1 ~ 1,
        control_numeric == 1 ~ 0
      )
  )

dat <- dat %>%
  mutate(
    Disorder =
      recode_factor(
        Disorder_numeric,
        `0` = "Controls",
        `1` = "Depressive disorders",
        `2` = "Anxiety disorders",
        `3` = "Depressive & anxiety disorders",
        `4` = "PTSD",
        `5` = "Eating disorders",
        `6` = "OCDs",
        `7` = "ADHD",
        `8` = "Personality disorders",
        `9` = "Autism",
        `10` = "Bipolar disorders",
        `11` = "Psychotic disorders"
      )
  )

dat %>%
  freq(Disorder)
```

```{r}
dat %>%
  filter(is.na(Disorder)) %>%
  freq(Sample)
```

@@

```{r gather to long format Disorder}
dat.long.psych <- dat %>%
  gather(
    key = "Disorder_inc_raw",
    value = "Diagnosis",
    all_of(disorder_columns)) %>%
  select(
    ID,
    Sample,
    Gender,
    Ethnicity,
    all_of(diff_score_cols),
    Disorder_inc_raw,
    Diagnosis
  ) %>%
  mutate(Disorder_inc =
           recode_factor(
             Disorder_inc_raw,
             anxiety_disorders_numeric = "Anxiety disorders",
             depressive_disorders_numeric = "Depressive disorders",
             eating_disorders_numeric = "Eating disorders",
             obsessive_compulsive_disorders_numeric = "OCDs",
             psychotic_disorders_numeric = "Psychotic disorders",
             mhd.mania_hypomania_bipolar_or_manicdepression_numeric = "Bipolar disorders",
             mhd.posttraumatic_stress_disorder_ptsd_numeric = "PTSD",
             mhd.autism_aspergers_or_autistic_spectrum_disorder_numeric = "Autism",
             mhd.attention_deficit_hyperactivity_disorder_numeric = "ADHD",
             mhd.personality_disorder_numeric = "Personality disorders",
             control_numeric = "Controls"
           ))

@@

```{r Disorder_inc by Gender}
Disorder_by_Gender <- dat.long.psych %>%
  filter(Diagnosis == 1) %>%
  count(Disorder_inc, Gender) %>%
  mutate(Prop = round(n/sum(n), 2))

Disorder_inc_by_Gender
```

```{r Disorder_inc by Gender}
Disorder_inc_by_Gender <- dat.long.psych %>%
  group_by(Disorder_inc) %>%
  filter(Diagnosis == 1) %>%
  count(Gender) %>%
  mutate(Prop = round(n/sum(n), 2))

Disorder_inc_by_Gender
```

```{r Disorder_inc by Gender plot}
Disorder_inc_by_Gender %>%
  ggplot(mapping = aes(
    x = Disorder_inc,
    y = n)
    ) +
  geom_bar(
    aes(fill = Gender),
#    position = "dodge",
    stat="identity",
    binwidth = 1,
    position = position_dodge2(preserve = "single")
  ) +
  labs(
    x = "Disorders",
    y = "Number of participants",
    fill = "Gender") +
  theme_personal +
  scale_fill_manual(
    values = COPINGpalette4,
#    labels = c("Prepandemic", "Retrospective")
    ) +
  coord_flip()
```

```{r Disorder_inc by Sample}
Disorder_inc_by_Sample <- dat.long.psych %>%
  filter(Diagnosis == 1) %>%
  count(Disorder_inc, Sample)
```

```{r Disorder_inc by Sample plot}
Disorder_inc_by_Sample %>%
  ggplot(mapping = aes(
    x = Disorder_inc,
    y = n)
    ) +
  geom_bar(
    aes(fill = Sample),
#    position = "dodge",
    stat="identity",
    binwidth = 1,
    position = position_dodge2(preserve = "single")
  ) +
  labs(
    x = "Disorders",
    y = "Number of participants",
    fill = "Sample") +
  theme_personal +
  scale_fill_manual(
    values = COPINGpalette4,
#    labels = c("Prepandemic", "Retrospective")
    ) +
  coord_flip()
```

Ethnicity should be recoded
European
Mixed or multiple ethnic origins
African or African British
Asian or Asian British
Arab
```{r Disorder_inc by Ethnicity}
Disorder_inc_by_Ethnicity <- dat.long.psych %>%
  filter(Diagnosis == 1) %>%
  count(Disorder_inc, Ethnicity) %>%
  mutate(Prop = round(n/sum(n), 2))

Disorder_inc_by_Ethnicity
```

```{r Disorder_inc by Ethnicity plot}
Disorder_inc_by_Ethnicity %>%
  ggplot(mapping = aes(
    x = Disorder_inc,
    y = n)
  ) +
  geom_bar(
    aes(fill = Ethnicity),
    #    position = "dodge",
    stat="identity",
    binwidth = 1,
    position = position_dodge2(preserve = "single")
  ) +
  labs(
    x = "Disorders",
    y = "Number of participants",
    fill = "Ethnicity") +
  theme_personal +
  scale_fill_manual(
    values = COPINGpalette6,
    #    labels = c("Prepandemic", "Retrospective")
  ) +
  coord_flip()
```

@@

```{r RUBBISH 1}
gad.diff_score_prepan_base_plot <- dat %>%
  ggplot(mapping = aes(y = gad.diff_score_prepan_base)) +
  geom_bar() +
  theme_minimal()

@@

```{r RUBBISH 2 fig.height=5, fig.width=5}
gad.diff_score_prepan_base_plot /
  gad.diff_score_prepan_retro_plot /
  gad.diff_score_prepan_base_plot +
  plot_layout(guides = "collect", heights = c(1,1,1)) +
  plot_annotation(tag_levels = "A")
```

@@

```{r RUBBISH }
#ADD
```

@@

Model 0: Disorders
```{r}
model0 <- lm(
  formula = gad.diff_score_prepan_retro ~ Disorder,
  data = dat)

@@

Model 1: Model 0 + Sex + Age + Ethnicity + Time registration and baseline
```{r}
model1 <- lm(
  formula = gad.diff_score_base_retro ~ Disorder + Gender + Age + Ethnicity + time_diff_sign_up_coping,
  data = dat)

@@

PCL:  
Calculate the total GLAD, COPING, RAMP baseline 
## pcl

```{r descriptives pcl.diff_score_cols}
pcl.numeric_table <- dat %>% 
#    group_by(Sample) %>%
  summarise_at( 
    vars(
      "Base - prepan" = pcl.diff_score_base_prepan
    ),
    .funs =
      list(
        ~mean(., na.rm = T),
        ~sd(., na.rm = T),
        ~median(., na.rm = T),
        ~min(., na.rm = T),
        ~max(., na.rm = T),
        ~skew(., na.rm = T),
        kurtosis = ~kurtosi(., na.rm = T)
      )
  )

pcl.numeric_table
```

```{r smart numeric summary tables}
# reshape the table
pcl.numeric_summary_table_reshaped <- pcl.numeric_table %>% 
  gather(key, value) %>% 
  separate(key, c("variable", "measure"), sep = "_") %>%
  mutate(measure = factor(
    x = measure,
    levels =  c( # this is the order I want my columns to appear in
      "missing",
      "min",
      "max",
      "mean", 
      "median", 
      "mode",
      "sd",
      "skew",
      "kurtosis"
    ))) %>% 
  spread(measure, value) %>% 
  # we can apply a function to all variables to edit them so that they are publication ready
  rename_all(
    .funs = ~str_to_title(.)
  ) %>% 
  # or we can individually rename variables 
  rename(
    SD = Sd
  )

pcl.numeric_summary_table_reshaped
```

Table for knitting
pcl diff summary table
```{r pcl diff summary table}
kable(pcl.numeric_summary_table_reshaped)
```

```{r descriptives pcl.diff_score_cols}
pcl.numeric_table <- dat %>% 
#    group_by(Sample) %>%
  summarise_at( 
    vars(
      "Prepan" = pcl.sum_score_prepan,
      "Base" = pcl.sum_score_base
    ),
    .funs =
      list(
        ~mean(., na.rm = T),
        ~sd(., na.rm = T),
        ~median(., na.rm = T),
        ~min(., na.rm = T),
        ~max(., na.rm = T),
        ~skew(., na.rm = T),
        kurtosis = ~kurtosi(., na.rm = T)
      )
  )

pcl.numeric_table
```

```{r smart numeric summary tables}
# reshape the table
pcl.numeric_summary_table_reshaped <- pcl.numeric_table %>% 
  gather(key, value) %>% 
  separate(key, c("variable", "measure"), sep = "_") %>%
  mutate(measure = factor(
    x = measure,
    levels =  c( # this is the order I want my columns to appear in
      "missing",
      "min",
      "max",
      "mean", 
      "median", 
      "mode",
      "sd",
      "skew",
      "kurtosis"
    ))) %>% 
  spread(measure, value) %>% 
  # we can apply a function to all variables to edit them so that they are publication ready
  rename_all(
    .funs = ~str_to_title(.)
  ) %>% 
  # or we can individually rename variables 
  rename(
    SD = Sd
  )

pcl.numeric_summary_table_reshaped
```

Table for knitting
pcl diff summary table
```{r pcl diff summary table}
kable(pcl.numeric_summary_table_reshaped)
```
